---
title: "World City Walkthrough"
author: "Brian"
date: "12/9/2021"
output: html_document
---

#Walkthrough Part 1: Plotly

```{r}
library(ggplot2)
library(tidyverse)
library(plotly)
library(leaflet)
library(sp)
library(flexdashboard)
library(shiny)
library(shinydashboard)
```


##random graph from scratch
```{r}
p <- plot_ly( 
  x = rnorm(1000),
  y = rnorm(1000),
  mode = 'markers'
)
```

```{r}
p #interactive plotly graph of randomly distributed numbers
```

##Scatter plot using plotly from scratch
```{r}
plot_ly(
  x = c(1,2,3,4,5,6,7),
  y = c(1,2,3,4,5,6,7),
  type = 'scatter',
  mode = 'markers'
) 
```

##Bar chart using plotly from scratch
```{r}
plot_ly(
  x = c(1,2,3),
  y = c(5,6,7),
  type = 'bar',
  mode = 'markers'
)
```

##Bubble Chart using plotly from scratch
```{r}
plot_ly(
  x = c(1,2,3),
  y = c(5,6,7),
  type = "scatter",
  mode = "markers",
  size = c(25,40,20),
  marker = list(color = c('red','blue','green')))
```

##Heatmaps using plotly from scratch
```{r}
plot_ly(
  z = volcano,
  type = "heatmap"
)
```

##Area Plot using plotly from scratch
```{r}
plot_ly(
  x = c(1,2,3),
  y = c(5,6,7),
  type = "scatter",
  mode = "lines",
  fill = "tozeroy"
)
```

```{r}
galton <- readRDS("galton.rds") #read in rds file
```


##Static Histogram using ggplot2
```{r}
ggplot(data = galton) + #static plot
  geom_histogram(aes(x = height, fill=sex), binwidth = 1, alpha = 0.5, position = "identity") +
  labs(
    title = "Child Height Vs Father Height (static)",
    x = "Father's Height (inches)",
    y = "frequency (n)"
  )
```


##Interactive Histogram using ggplotly
```{r}
ex1 <- ggplot(galton) + #dynamic plot
  geom_histogram(aes(x = height, fill=sex), binwidth = 1, alpha = 0.5, position = "identity") +
  labs(
    title = "Child Height Vs Father Height (dynamic)",
    x = "Father's Height (inches)",
    y = "frequency (n)"
  )
```


```{r}
ggplotly(ex1)
```

##Scatterplot using ggpplotly
```{r}
ggplot(galton) + #static plot
  geom_point(aes(father,height,color = sex)) +
  labs(
    title = "Child Height vs Father Height (static)",
    x = "Father Height (inches)",
    y = "Child Height (inches)"
  )
```


```{r}
ggplotly(ggplot(galton) + #dynamic plot
  geom_point(aes(father,height,color = sex)) +
  labs(
    title = "Child Height vs Father Height (dynamic)",
    x = "Father Height (inches)",
    y = "Child Height (inches)"
  ))
```

```{r}
ggplot(galton) + #static plot
  geom_point(aes(father, height, color = sex), size=2) +
  geom_smooth(aes(father,height), method="lm") +
  labs(
    title = "Child Height vs Father Height (static)",
    x = "Father's Height (inches)",
    y = "Child's Height (inches)"
  )
```

```{r}
ex2 <- ggplot(galton) + #static plot
  geom_point(aes(father, height, color = sex), size=2) +
  geom_smooth(aes(father,height), method="lm") +
  labs(
    title = "Child Height vs Father Height (static)",
    x = "Father's Height (inches)",
    y = "Child's Height (inches)"
  )
```

```{r}
ggplotly(ex2)
```


#LEAFLET - INTERACTIVE MAPS USING R
```{r}
map_example01 <- leaflet() %>% #quick example
  addTiles() %>% #default is OpenStreetMaps
  addMarkers(lng=174.786, lat = -36.852, popup = "Birthplace of R")
```

```{r}
map_example01 #The birthplace of R
```


#PREPPING YOUR FILES
```{r}
cities_file <- read.csv("worldcities.csv") #read in dataset
```

```{r}
head(cities_file) #examine data
```


```{r}
cities <- cities_file %>% #cleaning
  select(city_ascii,country,iso3, population,lat,lng) %>%
  rename(City = city_ascii,
         Country = country,
         Abbr = iso3
         )
```

#casting
```{r}
cities$lng <- as.numeric(cities$lng) #casting longitude as integer
cities$lat <- as.numeric(cities$lat) #casting latitude as integer
cities$population <- as.numeric(cities$population) #casting population as integer
```

```{r}
cities.sp <- SpatialPointsDataFrame(cities[,c(5,6)], cities[,-c(5,6)])
```


```{r}
cities_df <- data.frame(cities) #creates proper data frame
```



##Adding a background to your map
```{r}
bg_example1 = leaflet() %>% #add background
  addProviderTiles(providers$Esri.NatGeoWorldMap) #third party tiles
```

```{r}
bg_example1 #only has Nat Geo background
```

```{r}
bg_example2 <- leaflet() %>% #different background
  addProviderTiles("CartoDB.Positron")
```


```{r}
bg_example2 #CartoDB Background
```

```{r}
bg_example3 <- leaflet() %>% #different background
  addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
```

```{r}
bg_example3 #CartoDB Background
```

##Setting markers to your map
```{r}
mega_city <- cities %>% #subset of data of cities with population over 15 million
  filter(population > 8000000)
```


```{r}
example_markers <- leaflet() %>% #adding markers
  addTiles() %>% 
  addMarkers(data = mega_city, lng = ~lng, lat = ~lat, popup = ~City)
```

```{r}
example_markers
```



##CUSTOM POPUPS

```{r}
example_cmarkers1 <- leaflet() %>% #adding markers
  addTiles() %>% 
  addMarkers(data = mega_city, lng = ~lng, lat = ~lat, popup = ~paste("<h3>Name: ",City,"</h3><br>", sep=" "))
```                                                                          

```{r}
example_cmarkers1
```


```{r}
example_cmarkers2 <- leaflet() %>% #adding markers
  addTiles() %>% 
  addMarkers(data = mega_city, lng = ~lng, lat = ~lat, popup = ~paste("<h3>Name: ",City,"</h3><br>","Country: ", Country, "<br>Population: ",population, sep=" "))
```


```{r}
example_cmarkers2
```

```{r}
example_cmarkers3 <- leaflet() %>% #adding markers
  addProviderTiles("NASAGIBS.ViirsEarthAtNight2012") %>% 
  addMarkers(data = mega_city, lng = ~lng, lat = ~lat, popup = ~paste("<h3>Name: ",City,"</h3><br>","Country: ", Country, "<br>Population: ",population, sep=" "))
```

```{r}
example_cmarkers3
```

##SET VIEW
```{r}
example_view <- leaflet(mega_city) %>%
  addTiles() %>%
  setView(lng = 126.9, lat = 37.5, zoom = 6)
```

```{r}
example_view
```


##ADDING POLYLINES
```{r}
example_poly1 <- leaflet(mega_city) %>%
  addTiles() %>%
  addMarkers(data = mega_city, lng = ~lng, lat = ~lat, popup = ~City) %>%
  addPolylines(lat = ~lat, lng = ~lng, color = "green")
```


```{r}
example_poly1 #lines connecting the cities
```

```{r}
example_poly2 <- leaflet(mega_city) %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addPolylines(lat = ~lat, lng = ~lng, color = "green") %>%
  addPopups(lat = ~lat, lng = ~lng, popup = ~Abbr) #adds abbreviations
```


```{r}
example_poly2 #lines connecting the cities and adding abbreviations
```


